|

### 2021-09-10 article/article.Rmd
rewrite abstract, add changes to text

---
title:  |
        | The Effect of Legislature Size on Public Spending:
        | A Meta-Analysis
date: \today
abstract: "In a seminal article, @weingast1981political argue that there is a positive relationship between legislature size and inefficiency in public expenditures. Their proposition is currently known as the \"law of $1/n$\" and has been widely debated in political science and public administration. However, recent studies have questioned the validity of the theory. In this letter, we conduct the first meta-analysis that assesses the generality of the \"law of $1/n$\". Based on a sample of 30 articles, we find no robust evidence for the effect of legislature size on government budgets. Yet the aggregate results show significant heterogeneity. We find moderate support for the \"law of $1/n$\" in unicameral legislatures and in the upper house, but panel/fixed-effects models and regression discontinuity designs reduce public spending estimates. Electoral system does not strongly affect the \"law of $1/n$\", although one coefficient indicates that majoritarian systems may be more prone to overspending than others."
abstractspacing: double
keywords: distributive politics; law of $1/n$; legislature size; meta-analysis; public spending
jelcodes: H21; H23; H50; H61
fontsize: 11pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

@@

In this letter, we conduct the first meta-analysis that tests the generality of
the "law of $1/n$". We select `r length(unique(dat$id))` articles that
use quantitative methods to evaluate the impact of legislature size over
government spending across several dimensions. Our study sample mirrors the
diversity of the literature. We found studies that present both negative and positive coefficients, often reaching statistical significance in either. Given the volume and the disparity of
the studies, we employ meta-analysis to summarise the results. Meta-analysis
provides a rigorous approach to combine heterogeneous outcomes into a single
estimation, and it allows scholars to gain valuable insights from the
aggregated data [@cooper2019handbook; @hedges1985statistical]. Meta-analysis
can also identify potential sources of study variability, enabling researchers
to assess threats to external validity and direct future efforts into more
promising areas of academic inquiry [@doucouliagos2008democracy]. Research
synthesis methods have been successfully applied in medicine and psychology
since the 1970s [@glass2015meta], and our work contributes to the burgeoning
literature that uses meta-analytic methods to understand challenging questions
in political science [@costa2017responsive; @doucouliagos2008democracy;
@green2013field; @lau2007effects; @schwarz2020supporting].

We run meta-regressions to further evaluate the relationship
between legislature size and government expenditures. In those analyses, we
measure the effect of five moderators that capture different sources of
heterogeneity in the literature. In section G of the Supplementary Materials, we use binomial tests to assess whether the proportions of positive and negative coefficients in our
sample are statistically different from each other. We also construct a series of funnel plots to
test for publication bias in existing research on the "law of $1/n$"
[@easterbrook1991publication; @gerber2001testing].

@@

The third and fourth columns use the natural logarithm of expenditure per
capita as the dependent variable. Among the coefficients in the restricted
sample, those in published studies tend to be smaller than those in
working papers\footnote{We find no evidence of publication bias in our models.
The funnel plots for all estimations are available Sections H and I of the
Supplementary Material.}. Another two moderators are negatively associated with
the outcome in our larger coefficient pool. They both refer to estimation methods.
Studies that employ panel/fixed effects or regression discontinuity designs
(RDDs) have lower coefficients for log expenditure per capita if we take OLS as
the reference category.

Three estimates are statistically significant in the third set of
meta-regressions, which include public expenditure as a percentage of GDP as
the dependent variable. Both in our restricted and in our extended samples,
recent studies have smaller coefficients than older papers, which stands in
contrast with the first model. Moreover, institutional design also affects
outcomes. Papers that use mixed samples of bicameral and unicameral systems
report significantly lower coefficients than those that analyse bicameral
legislatures exclusively. Non-majoritarian electoral systems, a category that encompasses proportional representation and mixed systems, may have a positive effect on the "law of $1/n$", but the estimate only borders on statistical significance.

>>>>

---
title:  |
        | The Effect of Legislature Size on Public Spending:
        | A Meta-Analysis
date: \today
abstract: "In a seminal article, @weingast1981political argue that there is a positive relationship between legislature size and inefficiency in public expenditures. Their proposition is currently known as the \"law of $1/n$\" and has been widely debated in political science and public administration. However, recent studies have questioned the validity of the theory. In this letter, we conduct the first meta-analysis that assesses the generality of the \"law of $1/n$\". Based on a sample of 30 articles, we find no robust evidence suggesting that legislature size has either a positive or a negative effect on government budgets. Yet the aggregate results mask considerable heterogeneity. Our findings provide moderate support for the \"law of $1/n$\" in unicameral legislatures and in upper houses, but they also indicate that papers using panel/fixed-effects models or regression discontinuity designs report negative public spending estimates. We find only limited evidence that electoral systems impact public spending, what suggests that proportional representation systems may not be more prone to overspending than majoritarian ones."
abstractspacing: double
keywords: distributive politics; law of $1/n$; legislature size; meta-analysis; public spending
jelcodes: H21; H23; H50; H61
fontsize: 11pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

@@

In this letter, we conduct the first meta-analysis that tests the generality of
the "law of $1/n$". We select `r length(unique(dat$id))` articles that use
quantitative methods to evaluate the impact of legislature size over government
spending across several dimensions. Our study sample mirrors the diversity of
the literature. We found articles that present a positive association between
the number of legislators and public expenditures, others suggesting that such
relationship is negative, and yet others that claim that there is no
correlation between them. Given the volume and the disparity of the studies, we
employ meta-analysis to summarise the results. Meta-analysis provides a
rigorous approach to combine heterogeneous outcomes into a single estimation,
and it allows scholars to gain valuable insights from the aggregated data
[@cooper2019handbook; @hedges1985statistical]. Meta-analysis can also identify
potential sources of study variability, enabling researchers to assess threats
to external validity and direct future efforts into more promising areas of
academic inquiry [@doucouliagos2008democracy]. Research synthesis methods have
been successfully applied in medicine and psychology since the 1970s
[@glass2015meta], and our work contributes to the burgeoning literature that
uses meta-analytic methods to understand challenging questions in political
science [@costa2017responsive; @doucouliagos2008democracy; @green2013field;
@lau2007effects; @schwarz2020supporting].

We run meta-regressions to further evaluate the relationship between
legislature size and government expenditures. In those analyses, we measure the
effect of five moderators that capture different sources of heterogeneity in
the literature. In our Supplementary Materials, we also include binomial
tests to assess whether the proportions of positive and negative coefficients
are statistically different from each other, as well as funnel plots to test
for publication bias in existing research on the "law of $1/n$"
[@easterbrook1991publication; @gerber2001testing].

@@

The third and fourth columns use the natural logarithm of expenditure per
capita as the dependent variable. Among the coefficients in the restricted
sample, those in published studies tend to be smaller than those in
working papers\footnote{We find no evidence of publication bias in our models.
The funnel plots for all estimations are available Sections H and I of the
Supplementary Material.}. Two other moderators are negatively associated with
the outcome in our larger coefficient pool. They both refer to estimation methods.
Studies that employ panel/fixed effects or regression discontinuity designs
(RDDs) have lower coefficients for log expenditure per capita if we take OLS as
the reference category.

Three estimates are statistically significant in the third set of
meta-regressions, which include public expenditure as a percentage of GDP as
the dependent variable. Both in our restricted and in our extended samples,
recent studies have smaller coefficients than older papers, which stands in
contrast with the first model. Institutional design also affects outcomes.
Papers that include mixed political systems report significantly lower
coefficients than those that analyse bicameral exclusively. Non-majoritarian
electoral systems have a small, positive effect in our extended sample model,
yet the coefficient is only significant at the 10\% level.
